---
title: "DA 460 - Lab 3"
author: "Marjorie Blanco"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Handout 3 R: The normal distribution

## The Data

```{r}
download.file("http://www.openintro.org/stat/data/bdims.RData", destfile = "bdims.RData")
load("bdims.RData")

# view data
head(bdims)

# subset of record where sex is male
mdims <- subset(bdims, sex == 1)
# subset of record where sex is female
fdims <- subset(bdims, sex == 0)
```

### Exercise 1

Make a histogram of `men's heights` and a histogram of `women's heights`. How would you compare the various aspects of the two distributions?

Both distribution for men's height and women's height are nearly normally distributed.  The standard deviation for women's height is smaller than men's height.

## The normal distribution

### Men's heights

```{r}
# calculate mean and standard deviation for men's heights
mhgtmean <- mean(mdims$hgt)
mhgtsd   <- sd(mdims$hgt)
mhgtmedian   <- median(mdims$hgt)
freq <- tapply(mdims$hgt,mdims$hgt,length)
mhgtmode <- as.numeric(names(freq)[which.max(freq)])

# frequency histogram and density histogram for women's heights
hist(mdims$hgt, probability = TRUE, main="Histogram of Height for males", xlab = "Height", breaks = 20)
x <- min(mdims$hgt):max(mdims$hgt)
y <- dnorm(x = x, mean = mhgtmean, sd = mhgtsd)
lines(x = x, y = y, col = "blue")
```

The data follow a nearly normal distribution.

- Mean: `r mhgtmean`

- Median: `r mhgtmedian`

- Mode: `r mhgtmode`

- SD: `r mhgtsd`

### Women's heights

```{r}
# calculate mean and standard deviation for women's heights
fhgtmean <- mean(fdims$hgt)
fhgtsd   <- sd(fdims$hgt)
fhgtmedian <- median(fdims$hgt)
freq <- tapply(fdims$hgt,fdims$hgt,length)
fhgtmode <- as.numeric(names(freq)[which.max(freq)])

# frequency histogram and density histogram for women's heights
hist(fdims$hgt, probability = TRUE, main="Histogram of Height for females", xlab = "Height", breaks = 20)
x <- min(fdims$hgt):max(fdims$hgt)
y <- dnorm(x = x, mean = fhgtmean, sd = fhgtsd)
lines(x = x, y = y, col = "blue")
```

The data follow a nearly normal distribution.

- Mean: `r fhgtmean`

- Mean: `r fhgtmedian`

- Mode: `r fhgtmode`

- SD: `r fhgtsd`

The mean for women's height is smaller than the men's height.

### Exercise 2

Based on the this plot, does it appear that the data follow a nearly normal distribution?

The data for both women's height and men's height appears to follow a nearly normal distribution.  The mean, median and mode are not the same.

## Evaluating the normal distribution

```{r}
# construct a normal probability plot for women's height
qqnorm(fdims$hgt)
qqline(fdims$hgt)

# construct a normal probability plot for men's height
qqnorm(mdims$hgt)
qqline(mdims$hgt)
```

### Exercise 3

Make a normal probability plot of `sim_norm`. Do all of the points fall on the line? How does this plot compare to the probability plot for the real data?

Not all of the points fall on the line.  The majority of points fall on the line and the ones that don't are on the tail.

```{r}
# simulate data from a normal distribution for women's heights
sim_norm <- rnorm(n = length(fdims$hgt), mean = fhgtmean, sd = fhgtsd)
# construct a normal probability plot
qqnorm(sim_norm)
qqline(sim_norm)
```

Not all of the points fall on the line.  The majority of points fall on the line and the ones that don't are on the tail.

```{r}
# simulate data from a normal distribution for men's heights
sim_norm <- rnorm(n = length(mdims$hgt), mean = mhgtmean, sd = mhgtsd)
# construct a normal probability plot
qqnorm(sim_norm)
qqline(sim_norm)
```

### Exercise 4

Does the normal probability plot for `fdims$hgt` look similar to the plots created for the simulated data? That is, do plots provide evidence that the female heights are nearly normal?

The normal probability plot for `fdims$hgt` look similar to the simulated data plots.  The simulated data plots provide evidence that the female heights are nearly normal.

```{r}
# run QQplot simulation for female height
qqnormsim(fdims$hgt)
```

```{r}
# run QQplot simulation for male height
qqnormsim(mdims$hgt)
```

### Exercise 5

Using the same technique, determine whether or not female weights (`fdims$wgt`) appear to come from a normal distribution

The data does not appears to be normal distribution.  The normal probability plot for `fdims$wgt` does not look similar to the simulated data plots.  The simulated data plots provide evidence that the female weights are not normal distributed.  The data points for the female weight appear to have curvature.

```{r}
# calculate mean and standard deviation for women's weights
fwgtmean <- mean(fdims$wgt)
fwgtsd   <- sd(fdims$wgt)
fwgtmedian <- median(fdims$wgt)
freq <- tapply(fdims$wgt,fdims$wgt,length)
fwgtmode <- as.numeric(names(freq)[which.max(freq)])

# frequency histogram and density histogram for women's weights
hist(fdims$wgt, probability = TRUE, main="Histogram of Weight for females", xlab = "Weight")
x <- min(fdims$wgt):max(fdims$wgt)
y <- dnorm(x = x, mean = fwgtmean, sd = fwgtsd)
lines(x = x, y = y, col = "blue")
```

The data follow does not follow a normal distribution.

- Mean: `r fwgtmean`

- Median: `r fwgtmedian`

- Mode: `r fwgtmode`

- SD: `r fwgtsd`

Not all of the points fall on the line.  The some of points fall on the line.

```{r}
# construct a normal probability plot
qqnorm(fdims$wgt)
qqline(fdims$wgt)

# run QQplot simulation
qqnormsim(fdims$wgt)
```

## Normal probabilities

What is the probability that a randomly chosen young adult female is taller than 6 feet (about 182 cm)?

```{r}
# calculate probability by calculating a Z score and consulting a Z table, theoretical
1 - pnorm(q = 182, mean = fhgtmean, sd = fhgtsd)
# calculate the probability empirically
sum(fdims$hgt > 182) / length(fdims$hgt)
```

### Exercise 6

Write out two probability questions that you would like to answer; one regarding female heights and one regarding female weights. Calculate the those probabilities using both the theoretical normal distribution as well as the empirical distribution (four probabilities in all). Which variable, height or weight, had a closer agreement between the two methods?

Height had a closer agreement between the two methods.  The standard deviation for female height is `r fhgtsd`, while is `r fwgtsd` for female weight.

What is the probability that a randomly chosen young adult female is shoter than 5 feet (about 152 cm)?

```{r}
pnorm(q = 152, mean = fhgtmean, sd = fhgtsd)
sum(fdims$hgt < 152) / length(fdims$hgt)
```

The difference is `r pnorm(q = 152, mean = fhgtmean, sd = fhgtsd) - sum(fdims$hgt < 152) / length(fdims$hgt)`.

What is the probability that a randomly chosen young adult female is between 5 (about 152 cm) and 6 feet (about 182 cm)?

```{r}
pnorm(q = 182, mean = fhgtmean, sd = fhgtsd) - pnorm(q = 152, mean = fhgtmean, sd = fhgtsd)
sum(fdims$hgt > 152 & fdims$hgt < 182) / length(fdims$hgt)
```

The difference is `r pnorm(q = 182, mean = fhgtmean, sd = fhgtsd) - pnorm(q = 152, mean = fhgtmean, sd = fhgtsd) - sum(fdims$hgt > 152 & fdims$hgt < 182) / length(fdims$hgt)`.

What is the probability that a randomly chosen young adult female is taller than 160 cm?

```{r}
# calculate probability by calculating a Z score and consulting a Z table, theoretical
1 - pnorm(q = 160, mean = fhgtmean, sd = fhgtsd)
# calculate the probability empirically
sum(fdims$hgt > 160) / length(fdims$hgt)
```

The difference is `r 1 - pnorm(q = 160, mean = fhgtmean, sd = fhgtsd) - sum(fdims$hgt > 160) / length(fdims$hgt)`.

What is the probability that a randomly chosen young adult female is weights more than 60 kg?

```{r}
# calculate probability by calculating a Z score and consulting a Z table, theoretical
1 - pnorm(q = 60, mean = fwgtmean, sd = fwgtsd)
# calculate the probability empirically
sum(fdims$wgt > 60) / length(fdims$wgt)
```

The difference is `r 1 - pnorm(q = 60, mean = fwgtmean, sd = fwgtsd) - sum(fdims$wgt > 60) / length(fdims$wgt)`.

What is the probability that a randomly chosen young adult female is weights less than 50 kg?

```{r}
# calculate probability by calculating a Z score and consulting a Z table, theoretical
pnorm(q = 50, mean = fwgtmean, sd = fwgtsd)
# calculate the probability empirically
sum(fdims$wgt < 50) / length(fdims$wgt)
```

The difference is `r pnorm(q = 50, mean = fwgtmean, sd = fwgtsd) - sum(fdims$wgt < 50) / length(fdims$wgt)`.

What is the probability that a randomly chosen young adult female is between 55 kg and 80 kg?

```{r}
pnorm(q = 80, mean = fwgtmean, sd = fwgtsd) - pnorm(q = 55, mean = fwgtmean, sd = fwgtsd)
sum(fdims$wgt > 55 & fdims$wgt < 80) / length(fdims$wgt)
```

The difference is `r pnorm(q = 80, mean = fwgtmean, sd = fwgtsd) - pnorm(q = 55, mean = fwgtmean, sd = fwgtsd) - sum(fdims$wgt > 55 & fdims$wgt < 80) / length(fdims$wgt)`.