---
title: "DA 460 - Lab 3"
author: "Marjorie Blanco"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Handout 3 R: The normal distribution

## The Data

```{r}
download.file("http://www.openintro.org/stat/data/bdims.RData", destfile = "bdims.RData")
load("bdims.RData")

# view data
head(bdims)

# subset of record where sex is male
mdims <- subset(bdims, sex == 1)
# subset of record where sex is female
fdims <- subset(bdims, sex == 0)
```

### Exercise 1

Make a histogram of men's heights and a histogram of women's heights. How would you compare the various aspects of the two distributions?

## The normal distribution
```{r}
# calculate mean and standard deviation for men's heights
mhgtmean <- mean(mdims$hgt)
mhgtsd   <- sd(mdims$hgt)

# frequency histogram and density histogram for women's heights
hist(mdims$hgt, probability = TRUE, main="Histogram of Height for males", xlab = "Height", breaks = 20)
x <- min(mdims$hgt):max(mdims$hgt)
y <- dnorm(x = x, mean = mhgtmean, sd = mhgtsd)
lines(x = x, y = y, col = "blue")
```

The data follow a nearly normal distribution.

- Mean: `r mhgtmean`

- SD: `r mhgtsd`
```{r}
# calculate mean and standard deviation for women's heights
fhgtmean <- mean(fdims$hgt)
fhgtsd   <- sd(fdims$hgt)

# frequency histogram and density histogram for women's heights
hist(fdims$hgt, probability = TRUE, main="Histogram of Height for females", xlab = "Height", breaks = 20)
x <- min(fdims$hgt):max(fdims$hgt)
y <- dnorm(x = x, mean = fhgtmean, sd = fhgtsd)
lines(x = x, y = y, col = "blue")
```

The data follow a nearly normal distribution.

- Mean: `r fhgtmean`

- SD: `r fhgtsd`

The mean for women's height is smaller than the men's height.

### Exercise 2

Based on the this plot, does it appear that the data follow a nearly normal distribution?

The data appears to follow a nearly normal distribution.

## Evaluating the normal distribution

```{r}
# construct a normal probability plot
qqnorm(fdims$hgt)
qqline(fdims$hgt)
```

### Exercise 3

Make a normal probability plot of sim_norm. Do all of the points fall on the line? How does this plot compare to the probability plot for the real data?

Not all of the points fall on the line.  The majority of points fall on the line and the ones that don't are on the tail.

```{r}
# simulate data from a normal distribution for women's heights
sim_norm <- rnorm(n = length(fdims$hgt), mean = fhgtmean, sd = fhgtsd)
# construct a normal probability plot
qqnorm(sim_norm)
qqline(sim_norm)
```

### Exercise 4

Does the normal probability plot for fdims$hgt look similar to the plots created for the simulated data? That is, do plots provide evidence that the female heights are nearly normal?

The normal probability plot for fdims$hgt look similar to the simulated data plots.  The simulated data plots provide evidence that the female heights are nearly normal.

```{r}
# run QQplot simulation
qqnormsim(fdims$hgt)
```

### Exercise 5

Using the same technique, determine whether or not female weights appear to come from a normal distribution

The data does not appears to follow a nearly normal distribution.  The normal probability plot for fdims$wgt does not look similar to the simulated data plots.  The simulated data plots provide evidence that the female weights are not normal distributed.  The data points for the female weight appear to have curvature.

```{r}
# calculate mean and standard deviation for women's weights
fwgtmean <- mean(fdims$wgt)
fwgtsd   <- sd(fdims$wgt)

# frequency histogram and density histogram for women's weights
hist(fdims$wgt, probability = TRUE, main="Histogram of Weight", xlab = "Weight")
x <- min(fdims$hgt):max(fdims$hgt)
y <- dnorm(x = x, mean = fwgtmean, sd = fwgtsd)
lines(x = x, y = y, col = "blue")

# construct a normal probability plot
qqnorm(fdims$wgt)
qqline(fdims$wgt)

# run QQplot simulation
qqnormsim(fdims$wgt)
```

## Normal probabilities

What is the probability that a randomly chosen young adult female is taller than 6 feet (about 182 cm)?

```{r}
# calculate probability by calculating a Z score and consulting a Z table, theoretical
1 - pnorm(q = 182, mean = fhgtmean, sd = fhgtsd)
# calculate the probability empirically
sum(fdims$hgt > 182) / length(fdims$hgt)
```

### Exercise 6

Write out two probability questions that you would like to answer; one regarding female heights and one regarding female weights. Calculate the those probabilities using both the theoretical normal distribution as well as the empirical distribution (four probabilities in all). Which variable, height or weight, had a closer agreement between the two methods?

What is the probability that a randomly chosen young adult female is shoter than 5 feet (about 152 cm)?

```{r}
pnorm(q = 152, mean = fhgtmean, sd = fhgtsd)
sum(fdims$hgt < 152) / length(fdims$hgt)
```

What is the probability that a randomly chosen young adult female is between 5 (about 152 cm) and 6 feet (about 182 cm)?

```{r}
pnorm(q = 182, mean = fhgtmean, sd = fhgtsd) - pnorm(q = 152, mean = fhgtmean, sd = fhgtsd)
sum(fdims$hgt > 152 & fdims$hgt < 182) / length(fdims$hgt)
```

What is the probability that a randomly chosen young adult female is taller than 160 cm?

```{r}
# calculate probability by calculating a Z score and consulting a Z table, theoretical
1 - pnorm(q = 160, mean = fhgtmean, sd = fhgtsd)
# calculate the probability empirically
sum(fdims$hgt > 160) / length(fdims$hgt)
```

```{r}
# calculate probability by calculating a Z score and consulting a Z table, theoretical
1 - pnorm(q = 60, mean = fwgtmean, sd = fwgtsd)
# calculate the probability empirically
sum(fdims$wgt > 60) / length(fdims$wgt)
```