---
title: "Assignment 0"
author: "Marjorie Blanco"
date: "7/2/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(plyr)
library(ggplot2)
```

```{r, echo=FALSE}
# Read data source
source("http://www.openintro.org/stat/data/arbuthnot.R")
```

```{r}
# View data
arbuthnot

# View dimension
dim(arbuthnot)

# View names of these columns (or variables)
names(arbuthnot)

# View single column of a data
arbuthnot$boys
```

- Exercise 1: What command would you use to extract just the counts of girls baptized?

There are `r dim(arbuthnot)[1]` records for girls baptized in the data set.

```{r}
length(arbuthnot$girls)

dim(arbuthnot)[1]
```
- Exercise 2: Is there an apparent trend in the number of girls baptized over the years?
How would you describe it?

The is steady increase

```{r}
plot(x = arbuthnot$year, y = arbuthnot$girls, main="Number of girls baptized per year",
    xlab="Year", ylab="Number of girls baptized") 

plot(x = arbuthnot$year, y = arbuthnot$girls, type="l", main="Number of girls baptized per year",
    xlab="Year", ylab="Number of girls baptized")

ggplot(data = arbuthnot, aes(year,girls)) +
  geom_line()
```

```{r}
arbuthnot$boys + arbuthnot$girls

plot(x = arbuthnot$year, y = arbuthnot$boys + arbuthnot$girls, type="l", type="l", main="Number of girls and boys baptized per year",
    xlab="Year", ylab="Number of girls and boys baptized")

arbuthnot$boys / arbuthnot$girls

arbuthnot$boys / (arbuthnot$boys + arbuthnot$girls)

arbuthnot$boys > arbuthnot$girls
```

- Exercise 3: Now, make a plot of the proportion of boys over time. What do you see?

```{r}
plot(x = arbuthnot$year, y = arbuthnot$boys / (arbuthnot$boys + arbuthnot$girls), type="l", main="Proportion of boys baptized per year",
    xlab="Year", ylab="Proportion of boys baptized")
```

```{r, echo=FALSE}
# Read data source
source("http://www.openintro.org/stat/data/present.R")
```



1. What years are included in this data set? What are the dimensions of the data frame and what are the variable or column names?

There are a total of years included in this data set.  The dimensions are.  And the variable names are: `r names(present)`

```{r}
count(present$year)
```
2. How do these counts compare to Arbuthnot’s? Are they on a similar scale?
3. Make a plot that displays the boy-to-girl ratio for every year in the data set. What do
you see? Does Arbuthnot’s observation about boys being born in greater proportion
than girls hold up in the U.S.? Include the plot in your response.
4. In what year did we see the most total number of births in the U.S.? You can refer to the
help files or the R reference card http://cran.r-project.org/doc/contrib/Short- refcard.pdf to find helpful commands.

In `r present[which.max(present[,"total"]), "year"]` the most total number of births in the U.S. The total was `r present[which.max(present[,"total"]), "total"]

```{r}
present$total <- present$boys + present$girls
which.max( present[,"total"] )
```

```{r}
plot(x = present$year, y = present$boys + present$girls, type="l")
```
